---
title: "CovidCAR Package Overview"
date: "2023-05-18"
description: "CovidCAR is intended to facilitate Covid19 model building, ensembling, and evalutaion"
format:
  html:
    df-print: kable
    code-fold: show
    code-summary: "Hide code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    number-sections: false
    html-math-method: katex
    css: styles.css
    theme: flatly
    toc-location: left
    smooth-scroll: true
editor_options: 
  chunk_output_type: console
---
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 11pt;
}
pre {
  font-size: 11pt
}
</style>


    
# Overview
This introduction briefly outlines core functions used to preprocess observation data, build spatial-temporal models, and post-process model outputs.  Its purpose is to demonstrate a standard workflow not to provide an in depth examination of functions or model building techniques.  

## Resources  
+ [The CovidCAR GitHub page](https://github.com/JMHumphreys/CovidCAR)    
+ [CovidCAR Functions Reference](https://jmhumphreys.github.io/CovidCAR/reference/index.html)  
  
  
# Preliminaries   
**Load needed packages**  
A few of these packages are not available on CRAN and will need to be installed from other locations.  
```{r echo=TRUE, warning=FALSE, message=FALSE}
#comments and prompts
options(dplyr.summarise.inform = FALSE)
library(cli)


#wrangling
library(tidyverse)
library(lubridate)
library(arrow)
library(Hmisc)
library(yaml)

#spatial manipulation
library(sp)
library(sf)
library(spdep)
library(rgeos)
library(igraph)
library(maptools)
library(mapproj)

#census data
library(censusapi)

#forecast data
library(zoltr) #Not available on CRAN
#remotes::install_github("reichlab/zoltr")
library(covidHubUtils) #Not available on CRAN
#remotes::install_github("reichlab/covidHubUtils")

#inference
library(INLA) #Not available on CRAN
#install.packages("INLA",repos=c(getOption("repos"),
#INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)

library(EpiEstim)
library(forecast)
```

[**CovidCAR**](https://github.com/JMHumphreys/CovidCAR) currently on GitHub  
```{r}
library(CovidCAR)
#devtools::install_github("JMHumphreys/CovidCAR")
```

# Setup Analysis 
```{r}
2+2
```
